Universals of word order result from optimization of grammars for efficient communication

### 2019-03-04 second-draft/4-paper-names.tex
revising

Natural languages vary a lot in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  `letter', and a phrase translating to `to a friend'.
In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
While the two sentences containing words with the same meanings, the order of these words are entirely different in the two languages:
%However, the languages differ in the order in which these words and phrases are arranged: 
In Arabic, the verb stands at the beginning, followed by both the object and the adpositional phrase.
`To' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
In Japanese, the verb stands at the end; object and adpositional phrase precede it; `to' is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
%It turns out that these are deep and stable regularities:
It turns out that this variation reflects a deep and stable regularity:
While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
this is strongly correlated with the occurrence of pre- or postpositions:
Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
%However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
This correlation has an extremely strong empirical basis:
In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization. %, over 90 \% of languages satisfy this generalization .

@@

@@

@@

Theoretical studies of word order typology have proposed Dependency Length Minimization as a unifying principle behind Greenberg's correlations \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
On the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, it has been confirmed that languages reduce dependency length, though a direct link to Greenger's correlations has not been established beyond theoretical arguments.
This has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}.

@@

The sentences in Figure~\ref{fig:dep-trees} are annotated with \key{dependency trees}: these are directed trees describing the grammatical relations among words. For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} of the verb, rather than e.g. the subject or an indirect object. A dependency arc is drawn from a \key{head} (e.g. \emph{kaita} in Figure~\ref{fig:arabic-japanese}) to a \key{dependent} (e.g. \emph{tegami-o}). Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}. Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.

The dependency length of a sentence is the sum of the lengths of its syntactic dependencies \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}:
\begin{equation}
    \label{eq:depl}
\operatorname{DepL}(t; {\bf w}) = \sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
\end{equation}
where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$, and the sum is negative because dependency length is seen as a cost.

Large-scale corpus evidence across languages shows that natural languages order words in such a way as to reduce this quantity, compared to most random orderings \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
However, the link to Greenberg's correlations has not been established this directly.
While theoretical arguments suggest that DLM predicts Greenberg's correlations, this idea has not been tested on the basis of actual large-scale corpus data.

We test this idea by constructing languages that have minimal dependency length:
If Greenberg's correlations emerge in these languages, then this provides evidence that DLM gives rise to these universals.
All natural languages have some degree of word order regularities that are specified in their grammar.
For instance, English places subjects before and objects after the verb (Figure~\ref{fig:dep-trees}).
Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
This is illustrated in Figure~\ref{fig:dlm-grammars}: Grammars specify how dependency structures are encoded into strings of words.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.

We create optimized grammars that minimize the average dependency length.
Prior theoretical work on dependency length minimization has argued that certain general patterns minimize dependency length \cite{temperley2008dependency}, however, this depends on the topology of the dependency trees.
Answering these questions requires a sample of dependency structures as actually used by speakers of difefrent languages.
Such samples have recently become available with the Universal Dependencies project, which has collected and created dependency annotations for several dozens of different languages.
We use data from 51 languages.
These corpora represent a typologically and genetically diverse group of languages.
Previously, \cite{temperley2008dependency, CITE} have created grammars optimized for DLM; ours is the first study applying this to study correlation universals on a large-scale basis.

@@

@@

%adp. &
	adposition    &    NP       
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} \textbf{86}  }
	&   \multirow{2}{*}{  \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf}   \textbf{47 \%}$^{}$   }
	&   \multirow{2}{*}{  \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf} \textbf{68 \%}$^{***}$    } \\
	\emph{to}            & \emph{a friend} \\ \hline
copula    &    NP         
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf}}
	&   \multirow{2}{*}{  \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf}   }
	&   \multirow{2}{*}{  \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf}}  \\
	\emph{is}        & \emph{a friend} \\ \hline
auxiliary    &    VP       &   \\
	\emph{has}          & \emph{written} \\
noun    &    genitive      &  \\
	\emph{friend} &  \emph{of John} \\
noun    &    relative clause       &  \\
	\emph{books} & \emph{that you read} \\
complementizer    &    S        & \\
	\emph{that} & \emph{Mary has arrived} \\
verb    &    PP     &    \\
	\emph{went} & \emph{to school} \\
want    &    VP      &    \\
	\emph{wants}   &  \emph{to leave} \\
verb    &    subject    &  \\
	\emph{(there) entered} & \emph{a person} \\
verb    &    manner adverb     & \\
	\emph{ran} & \emph{quickly} \\
 \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
    \multicolumn{6}{l}{\footnotesize{Significance levels: $^*$: $p < 0.05$, $^{**}$: $p < 0.01$, $^{***}$: $p < 0.001$}}
\end{tabular}
	\end{center}
	\caption{Greenbergian Correlations. Following \cite{dryer1992greenbergian}, each correlation is stated in terms of a pair of a `verb patterner' and an `object patterner', whose relative order correlate with that of verbs and objects.
For each correlation, we provide an example.
Given the statistical nature of the correlations, not every language satisfies every one of them.
Not all correlations are satisfied by every natu
}\label{table:corr-dryer}
\end{table*}

\paragraph{Results}
We show that optimizing grammars for DLM produces Greenberg's correlations.
We base our evaluation on the authoritative compilation by Dryer \cite{dryer1992greenbergian}, which draws on 625 languages. %, which we take as the basis of our evaluation.
Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.

Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages, or are not annotated reliably in UD.
%One further dependency is not annotated reliably in UD. %: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations (See SI for information on the other correlations). % from Dryer's 15 correlations.
These are described in Table~\ref{table:corr-dryer}.
In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.

For each optimized grammar, we can check which correlations it satisfies.
%In order to test whether an objective function predicts a correlation, we selected all word order grammars created for the given function, and counted the percentage of grammars satisfying the correlation.
We conducted, for each correlation, a mixed-effects logistic regression model predicting whether a grammar satisfies the correlation, with random effects for languages and language families.\footnote{We coded language families according to \url{universaldependencies.org}.}
%We conducted, for each correlation, a mixed-effects logistic regression model predicting whether $a_\tau > 0.5$ holds for the correlating dependency depending on whether $a_\tau > 0.5$ for the verb-object dependency, with per-language-family random intercept and slope.\footnote{We coded language families according to \url{universaldependencies.org}.}
%We are interested in the direction and significance of this effect: 
%If the effect is significant, in the positive direction, we can conclude that a correlation is predicted across corpora from languages belonging to different language families.

@@

We have verified that dependency length minimization produces Greenberg's correlation universals.
However, dependency length is only a heuristic and stipulative measure of complexity: % heuristic and stipulative.
The actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}. 
It is also stipulative: There is no a priori reason why dependency length should be the crucial determining factor determining word order. There are other models of processing cost that make different predictions.
DLM is also insufficient as an underlying principle of word order, as absolutely minimizing dependency length results in languages that have no consistent grammatical orders, making comprehension difficult.
Indeed, theoretical work has justified DLM as describing an aspect of a broader notion of language efficiency.
\cite{hawkins1994performance} theoetically argues that DLM enables more efficient parsing.
\cite{futrell2017noisycontext} argues that DLM arises from the optimization of predictability.

@@

Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort (Force of Unification) and listener effort (Force of Diversification).
According to this theory, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
Conversely, to minimize listener effort, the language should provide different utterances for different meanings, so that the listeenr can unambiguously identify the meaning from the utterance.
Minimizing speaker and listener effort represent two opposing forces:
A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.
The idea that language results from the tension between these two pressures has a long and fruitful history.
%\cite{horn-toward-1984}, RSA, Color naming
%
%In the domain of word order, 
%
%TODO introduce the terms \key{cost} and \key{informativity}
%
This idea has been shown to account for phenomena such as pragmatic inference \cite{horn-toward-1984, frank2012predicting} and color naming~\cite{zaslavsky2018efficient}.

@@

We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
In Shannon's model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.
These are encoded in the dependency structure.
%Following a long tradition in formal and computational liguistics, we formalize these in the format of Dependency Grammar:
%This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
In Figure~\ref{fig:comm}, the message consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.

@@

Speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences
For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.

>>>>

Natural languages vary a lot in the order in which they express information.
Consider Figure~\ref{fig:arabic-japanese-simple}, showing a sentence in Arabic (top) and Japanese (bottom), both translating to `I wrote a letter to a friend.'
Both sentences contain a verb meaning `wrote', a noun expressing  `letter', and a phrase translating to `to a friend'.
In linguistic terminology, `letter' is the object of the verb, whereas the phrases translating as `to a friend' are known as adpositional phrases.
While the two sentences containing words with the same meanings, the order of these words are entirely different in the two languages:
%However, the languages differ in the order in which these words and phrases are arranged: 
In Arabic, the verb stands at the beginning, followed by both the object and the adpositional phrase.
`To' is expressed by a \emph{preposition}, so named because it precedes the word expressing `friend'. %The adpositional phrase starts with the adposition.
In Japanese, the verb stands at the end; object and adpositional phrase precede it; `to' is expressed by a \textit{postposition}, so named because it follows the word denoting `friend'.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb, and that the objects of adpositions and verbs stand on the sam side of their heads.
%What both languages have in common is that the object and the adpositional phrase stand on the same side of the verb.
%It turns out that these are deep and stable regularities:
It turns out that this variation reflects a deep and stable regularity:
While languages ordering the objects before (Japanese) or after (Arabic) the verb are approximately equally common around the world,
this is strongly correlated with the occurrence of pre- or postpositions:
This is illustrated in Figure~\ref{fig:corr-table}.
Languages ordering their objects the way Japanese does, have postpositions; languages ordering them as Arabic does have prepositions.
%However, the vast majority of natural languages place adpositional phrases on the same side of the verb as the object:
As Figure~\ref{fig:corr-table} shows, this correlation has an extremely strong empirical basis:
In a sample of about 1,100 languages from five contents (CITE Dryer WALS 95a), less than 7 \% of languages provide clear exceptions to this generalization. %, over 90 \% of languages satisfy this generalization .

\begin{figure}
\centering
\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
%	   I \& wrote \& the \& letter \& in \& Japanese \& . \\
	   katabt \& risāla \& li \& sadīq  \\
%	   \textsc{verb} \& \textsc{noun} \& \textsc{adp} \& \textsc{noun} \\
	   wrote \& letter   \& to \& friend  \\
   \end{deptext}
\end{dependency}
\begin{dependency}[theme = simple]
   \begin{deptext}[column sep=1em]
      tomodachi \& ni \& tegami-o \& kaita \\
 %\textsc{noun} \& \textsc{adp} \& \textsc{noun} \& \textsc{verb}\\
	   friend \& to \& letter  \& wrote \\
   \end{deptext}
\end{dependency}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{figures/corr-table.png}
	\caption{The idea of a word order correlation: Languages can order the object before or after the verb (rows), and have prepositions or postpositions (columns). For each combination, we indicate how many languages satisfy it, as documented in the World Atlas of Language Structure~\cite{wals}. Combinations on the diagonal are vastly more common than off-diagonal ones.}
	\label{fig:corr-table}
\end{figure}

@@

Theoretical studies of word order typology have proposed Dependency Length Minimization as a unifying principle behind Greenberg's correlations \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.
Our first contribution is to show that this principle indeed correctly predicts Greenberg's correlations.
We do this by creating counterfactual grammars that minimize dependency length of

@@

Theoretical studies of word order typology have proposed Dependency Length Minimization as a unifying principle behind Greenberg's correlations \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
This is the idea that languages order words so as to minimize the linear distance between words linked in head--dependent relationships.

@@

The sentences in Figure~\ref{fig:dep-trees} are annotated with \key{dependency trees}:
these are directed trees describing the grammatical relations among words.
For example, the arcs labeled ``obj'' represent that the noun in question is the \emph{direct object} of the verb, rather than e.g. the subject or an indirect object.
A dependency arc is drawn from a \key{head} (e.g. \emph{visited} in Figure~\ref{fig:dep-trees}) to a \key{dependent} (e.g. \emph{friends}). Dependency trees can be defined in terms of many different syntactic theories \cite{corbett1993heads}. Although there are some differences in how different formalisms would draw trees for certain sentences, there is broad enough agreement about dependency trees that it has been possible to develop large-scale dependency-annotated corpora of text from dozens of languages \cite{nivre2017universal}.

The \key{dependency length} of a sentence ${\bf w}$ with dependency structure $t$ is the sum of the lengths of its dependency arcs, measured as the number of words spanned \cite{ferrericancho2004euclidean,liu2008dependency,gildea2010grammars,futrell2015largescale}.
%\begin{equation}
%    \label{eq:depl}
%\operatorname{DepL}(t; {\bf w}) = \sum_{i=1}^{\# {\bf w}} |i - head(t; w_i)|,
%\end{equation}
%where $head(t; i) \in \{1, ..., n\}$ is the index of the head of $w_i$. 
%On the basis of large-scale corpus evidence \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}, it has been confirmed that languages reduce dependency length, though a direct link to Greenger's correlations has not been established beyond theoretical arguments.
Large-scale corpus evidence across languages shows that natural languages order words in such a way as to reduce dependency length, compared to most random orderings \cite{futrell2015largescale,liu2017dependency,temperley2018minimizing}.
These data provide strong evidence that minimization of dependency length is itself a linguistic universal~\cite{temperley2018minimizing}

While theoretical arguments suggest that DLM predicts Greenberg's correlations, this idea has not been tested on the basis of actual large-scale corpus data.
Prior theoretical work on dependency length minimization has argued that Greenberg's correlations minimize dependency length \cite{temperley2008dependency}; however, the link is not trivial:
Whether or not Greenberg's correlations theoretically minimize dependency length will depend on the distribution of tree topologies in actual language usage. %this depends on the topology of the dependency trees.

We test this idea by constructing languages that have minimal dependency length:
That is, we ask: Under the assumption that language has evolved to minimize dependency length, will Greenberg's correlations emerge?
We computationally construct hypothetical variants of real languages that are optimized to minimize dependency length.
If Greenberg's correlations emerge in these languages, then this provides evidence that DLM gives rise to these universals.

All natural languages have some degree of word order regularities that are specified in their grammar.
For instance, English places subjects before and objects after the verb (Figure~\ref{fig:dep-trees}).
Natural languages differ in the rules they apply: Some place the object after the verb, some place it before the verb.
This is illustrated in Figure~\ref{fig:dlm-grammars}: Grammars specify how dependency structures are encoded into strings of words.
For instance, Grammar 1 -- corresponding to Arabic in Figure~\ref{fig:arabic-japanese-simple} -- orders objects (`friends', `letter') after verbs and has prepositions (`to friend').
Grammar 2 orders objects after verbs but has postpositions (`friend -- to').
Grammars 3 and 4 place the object before the verb, and one of them (Grammar 3) corresponds to Japanese order.

We create optimized grammars that minimize the average dependency length.
Prior theoretical work on dependency length minimization has argued that certain general patterns minimize dependency length \cite{temperley2008dependency}, however, this depends on the topology of the dependency trees.
Answering these questions requires a sample of dependency structures as actually used by speakers of difefrent languages.
Such samples have recently become available with the Universal Dependencies (UD) project, which has collected and created dependency annotations for several dozens of different languages.
We use data from 51 languages.
These corpora represent a typologically and genetically diverse group of languages.
Previously, \cite{temperley2008dependency, CITE} have created grammars optimized for DLM; ours is the first study applying this to study correlation universals on a large-scale basis.

@@

%adposition    &    NP       &    86    &    \textbf{81}$^{***}$    &    47    &    \textbf{76}$^{***}$    &    \textbf{68}$^{***}$   \\
%copula    &    NP        &    94    &    \textbf{81}$^{***}$    &    53    &    \textbf{79}$^{***}$    &    \textbf{61}$^{**}$   \\
%auxiliary    &    VP       &    88    &    \textbf{74}$^{***}$    &    \textbf{84}$^{***}$    &    55    &    \textbf{69}$^{**}$   \\
%noun    &    genitive      &    80    &    \textbf{82}$^{***}$    &    55    &    \textbf{74}$^{***}$    &    \textbf{70}$^{***}$   \\
%noun    &    relative clause       &    80    &    \textbf{85}$^{***}$    &    48    &    \textbf{77}$^{**}$    &    \textbf{73}$^{***}$   \\
%complementizer    &    S        &    76    &    \textbf{85}$^{***}$    &    \textbf{59}$^{**}$    &    \textbf{80}$^{***}$    &    \textbf{74}$^{**}$   \\
%verb    &    PP     &    88    &    \textbf{78}$^{***}$    &    \textbf{72}$^{***}$    &    59    &    \textbf{69}$^{**}$   \\
%want    &    VP      &    88    &    \textbf{90}$^{***}$    &    \textbf{78}$^{**}$    &    \textbf{92}$^{***}$    &    \textbf{92}$^{***}$   \\
%verb    &    subject    &    33    &    \textbf{29}$^{**}$    &    51    &    \textbf{8}$^{***}$    &    \textbf{13}$^{***}$   \\
%verb    &    manner adverb     &    35    &    51    &    \textbf{21}$^{***}$    &    51    &    \textbf{32}$^{***}$   \\
%

@@

	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_case.pdf} \textbf{86}  }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_case.pdf}   \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_case.pdf}   \textbf{47 \%}$^{}$   }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_case.pdf}  \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_case.pdf} \textbf{68 \%}$^{***}$    } \\
	\emph{to}            & \emph{a friend} \\ \hline
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_cop.pdf} \textbf{94} }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_cop.pdf}   \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_cop.pdf}  \textbf{53}     }
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_cop.pdf}  \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_cop.pdf} \textbf{61}$^{***}$  }  \\
	\emph{is}        & \emph{a friend} \\ \hline
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-aux.pdf}}
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-aux.pdf}   \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-aux.pdf}   }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-aux.pdf}   \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-aux.pdf}}  \\
	\emph{has}          & \emph{written} \\ \hline
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-nmod.pdf}}
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-nmod.pdf}   }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-nmod.pdf}}  \\
	\emph{friend} &  \emph{of John} \\ \hline
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-acl.pdf}}
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-acl.pdf}\textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-acl.pdf}   }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-acl.pdf} \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-acl.pdf}}  \\
	\emph{books} & \emph{that you read} \\ \hline
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-lifted_mark.pdf}}
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-lifted_mark.pdf} \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-lifted_mark.pdf}   }
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-lifted_mark.pdf}\textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-lifted_mark.pdf}}  \\
	\emph{that} & \emph{Mary has arrived} \\ \hline
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-obl.pdf}}
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-obl.pdf} \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-obl.pdf}   }
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-obl.pdf}\textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-obl.pdf}}  \\
	\emph{went} & \emph{to school} \\ \hline
	&  \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-ground-xcomp.pdf}}
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-depl-nmod.pdf} \textbf{81}$^{***}$    }
	&    \multirow{2}{*}{\includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-langmod-xcomp.pdf}   }
	&   \multirow{2}{*}{ \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-parser-nmod.pdf} \textbf{76}$^{***}$ }   
	&   \multirow{2}{*}{  \includegraphics[width=0.03\textwidth]{../results/correlations/figures/correlations/correlation-efficiency-xcomp.pdf}}  \\
	\emph{wants}   &  \emph{to leave} \\ \hline
%comp. &
%\multirow{3}{*}{NP}&
%	AP &
For each correlation, we provide an example.
Given the statistical nature of the correlations, not every language satisfies every one of them.
Not all correlations are satisfied by every natu
}\label{table:corr-dryer}

\paragraph{Results}
We show that optimizing grammars for DLM produces Greenberg's correlations.
%Dryer \cite{dryer1992greenbergian} provides an authoritative
We base our evaluation on the authoritative compilation of Greenberg's correlations by Dryer \cite{dryer1992greenbergian}, which documents 15 correlations, drawing on 625 languages. %, which we take as the basis of our evaluation.
%Most of them can be straightforwardly implemented in the UD formalism, allowing us to check which correlations a word order grammar satisfies.
%Considering those correlations that are annotated reliably in UD, we obtain eight correlations
We exclude three correlations that are only relevant to some of the world's languages, and collapse correlations that are not distinguished in UD.
%Dryer \cite{dryer1992greenbergian} presents four correlations that are only relevant to part of the world's languages, or are not annotated reliably in UD.
%One further dependency is not annotated reliably in UD. %: the dependencies between question particles and verbs, those between nouns and plural words, and those between nouns and articles.
%Further collapsing correlations that cannot be formalized individually in UD, we otained eight formalized correlations (See SI for information on the other correlations). % from Dryer's 15 correlations.
We describe the resulting eight correlations in Table~\ref{table:corr-dryer}.
%In \cite{dryer1992greenbergian}, all word order correlations are relative to the position of the direct object wrt the main verb of a sentence.

For each optimized grammar, we can check which of these correlations it satisfies.
We conducted a multivariate mixed-effects logistic regression model predicting, for each grammar, which of the eight correlations it satisfies.
We entered the language and language family of the underlying corpus as random effects, accounting for language-specific properties of tree topologies.\footnote{We coded language families according to \url{universaldependencies.org}.}

We compare the prevalence of the word order correlations in simulated languages to their prevalence in the real languages.
To evaluate their presence in real languages, we tested for the correlations in word order grammars fit by maximum likelihood to actual orderings from treebanks.
The word order correlations detected this way match linguistic descriptions compiled in the World Atlas of Linguistic Structures (WALS, \cite{wals}) to the extent that they are documented in WALS. %in all except four languages, for WALS did not have data.

@@

We have verified that dependency length minimization produces Greenberg's correlation universals.
However, dependency length is only a heuristic and stipulative measure of complexity: % heuristic and stipulative.
The actual processing complexity induced by long dependencies is not a linear function of length and depends crucially on the types of dependencies involved \cite{demberg2008data} and the specific elements intervening between the head and dependent \cite{gibson1998linguistic,gibson2000dependency,lewis2005activationbased}. 
It is also stipulative: There is no a priori reason why dependency length should be the crucial determining factor determining word order.
While DLM has been justified based on the reasoning that long dependencies create high memory requirements for parsing and generation \cite{gibson1998linguistic,gibson2000dependency}, there are other models of processing cost that make different predictions.
In particular, surprisal is known to be a strong predictor of human processing effort, but it makes predictions different from and sometimes opposite to those made by dependency length.
DLM is also insufficient as an underlying principle of word order, as absolutely minimizing dependency length results in languages that have no consistent grammatical orders, making comprehension difficult.
Indeed, theoretical work has justified DLM as describing an aspect of a broader notion of language efficiency.
\cite{hawkins1994performance} theoetically argues that DLM enables more efficient parsing.
\cite{futrell2017noisycontext} argues that DLM arises from the optimization of predictability.

@@

Zipf \cite{zipf1949human} argued that language optimizes a tradeoff between speaker effort, the \emph{Force of Unification}, and listener effort, the \emph{Force of Diversification}.
According to this theory, a language that minimizes speaker effort should reduce the number of different utterances to make production as easy as possible.
Conversely, to minimize listener effort, the language should provide different utterances for different meanings, so that the listeenr can unambiguously identify the meaning from the utterance.
Minimizing speaker and listener effort represent two opposing forces:
A pressure to reduce the complexity of the language makes the utterances in a language more homogenous.
A pressure to reduce ambiguity makes utterances more heterogeneous, so they can indicate different meanings.
%Another way to conceptualize these two pressures is as the \key{cost} and \key{informativity} of language use:
Producing and processing simple utterances incurs little cost, but more complex and diverse utterances are required to be provide higher informativity.

The idea that language results from the tension between these two pressures has a long and fruitful history.
%\cite{horn-toward-1984}, RSA, Color naming
%
%In the domain of word order, 
%
%TODO introduce the terms \key{cost} and \key{informativity}
%
This idea has been shown to provide a formalization and derivation of pragmatic inference \cite{horn-toward-1984, frank2012predicting}.
It also has been shown to account for the distribution of word meanings across languages, such as color naming~\cite{zaslavsky2018efficient}.

@@

We model the process where a speaker communicates a message to a listener (Figure~\ref{fig:comm}) in Shannon's framework of information theory.
In Shannon's model, a transmitter encodes a message into a signal. The receiver decodes the signal, attempting to reconstruct the original message.
Applying this model to word order, we take the message to consist of entities, events, and a set of syntactic and semantic relations that hold between them.

Dependency structures provide a natural formalization of such structures (CITE).
%These are encoded in the dependency structure.
%This linguistic formalism draws directed arcs between syntactically related words, annotated with syntactic relations.
In Figure~\ref{fig:comm}, the message consists of a dependency structure with the words `saw', `Mary', `John', where Mary is the subject of the event denoted by `saw', and John is the object.

@@

Speakers and listeners have shared knowledge of a \key{grammar} that specifies how dependency structures are encoded into sentences.
For instance, English places subjects before and objects after the verb, allowing the listener to unambiguously decode the sentence in Figure~\ref{fig:comm}.
